---
title: "Adding custom CSS styles to a module"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adding custom CSS styles to a module}
  %\VignetteEngine{knitr::rmarkdown}14:55
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = "styler"
)
```

## Adding custom CSS styles to a module

There are multiple ways of adding custom CSS styles to a Shiny module:

1.  Inline styles

a. insert the style at the tag level, through the style attribute

```{r, eval=FALSE}
p(style = "color:red;", "Red text")
```

b. insert the style at the `head` level, through the style tag

```{r, eval=FALSE}
ui <- fluidPage(
  tags$head(
    tags$style(HTML("
      .shiny-input-container {
        color: #474747;
      }"))
  )
)
```

2.  As an external css file

a. Insert the style in the `head` of the HTML page

```{r, eval=FALSE}
ui <- fluidPage(
  tags$head(
    tags$link(rel = "stylesheet", type = "text/css", href = "custom_file.css")
  )
)
```

b. Insert the style via `includeCSS()` function call

```{r, eval=FALSE}
ui <- fluidPage(
  includeCSS("www/custom_file.css")
)
```

These are **general** options and can be used in all Shiny applications. See [Using custom CSS in your app](https://shiny.rstudio.com/articles/css.html) and [Style your apps with CSS](http://shiny.rstudio-staging.com/articles/css.html) for more examples and in-depth explanations.

However, for the purposes of working with **dv.manager** we recommend using a 3rd alternative in order to avoid `CSS leaking` (your specific CSS style being applied to other parts/modules of a larger Shiny application unintentionally).

### Add custom CSS reference as an HTML dependency

1. Define your CSS styles in a separate file and place it inside the `www` folder of your package

2. Inside the UI of your Shiny module, define an HTML dependency pointing to that file

```{r, eval=FALSE}
ui <- function(id) {
  dependency <- htmltools::htmlDependency(
    name = "custom-package_name_dep", # use a name that includes your package name,
    # so it will be easier to debug later
    version = 1, # for our purposes any value will suffice
    src = "/path/to/custom_css_file/",
    stylesheet = "custom.css",
    package = "package_name" # needed when the src parameter is relative
  )
  dependency
}
```

For more options, please check `htmltools::htmlDependency()` documentation page.

3. Attach the dependency variable to the UI structure. For example:

```{r, eval=FALSE}
ui <- function(id) {
  dependency <- htmltools::htmlDependency(
    # ...
  )

  shiny::tagList(
    # ...,
    dependency
  )
}
```

For other options of attaching an HTML dependency to a UI element, please consult `htmltools::attachDependencies` documentation.

### CSS namespacing in module manager

Advantages:

- Custom CSS module dependencies impact only that module.
- The `dv.manager` can still change the styles of all other modules if desired (for example with a custom Shiny theme)
- However the modules cannot impact the `dv.manager` CSS styles even with the use of `!important` keyword

Limitation:

- This approach only prevents **unintentional** CSS style overflow. It cannot work against JavaScript libraries (like the ones contained in styling packages: `shinydashboard`, `flexdashboard` and similar) which **actively** and **purposefully ** change and overwrite class names and CSS styles at runtime.
- Dependency name must begin with the prefix **"custom-"** otherwise it will not be namespaced.
