---
title: "Data reloading"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data reloading}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = "styler"
)
```

**dv.manager** allows periodical reloading of the datasets.

In the `run_app()` call we specify the duration after which the datasets should be reloaded via the `reload_period` parameter.

```{r, eval=FALSE}
dv.manager::run_app(
  data = list("DS" = data),
  module_list = module_list,
  reload_period = lubridate::duration(12, "hours")
)
```

The `reload_data` parameter is specified as either a lubridate object to specify a duration or a positive numeric value which then is interpreted as a lubridate duration object in days.  
By default the **dv.manager** does not reload the datasets after a specific time, as the `reload_data` parameter is NULL.

The functionality to reload the datasets has been implemented via `"touching"` an `"restart.txt"` file and therefore the `"shiny-server"` will then cover the reloading of the application and therefore a reloading of the datasets.

