---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = "styler"
)
```

Here, we will follow some simple steps to create an application using **dv.manager**. Note that that **dv.manager** does not include any module it just serves as a container therefore along these steps we will install the *app* package from the DaVinci project that contains several exploratory modules.

## Data loading

**dv.manager** has been developed to work with DaVinci modules whose main purpose is to explore clinical trial data. The main input for this modules is data loaded from SDTM or ADAM datasets. We will use two dummy datasets in the dv.manager for illustration purposes.

```{r, eval=FALSE}
data <- list(adsl = pharmaverseadam::adsl, adae = pharmaverseadam::adae)
```

**Notice** that datasets are all packed together in a single list.

## Preprocessing the data (optional)

We can optionally preprocess the loaded data by: computing derived variables, rename variables, select a subset of the variables or a subset of the population,...

## Module selection

Now, we have to select which modules will be included in our application. As mentioned before, **dv.manager** does not contain any modules, therefore we will install the *dv.general* package (see how to install *dv.general* an other modules in: `vignette("installing_module_packages")`).

If you already have **dv.general** installed you can skip this step.

```{r, eval=FALSE}
# TODO
repos <- options("repos")

install.packages("dv.general", repos = repos)
```

We can now include in our application or the modules included in the **dv.general** package.

For example, we can use the *data table* from the **dv.general** package.

```{r, eval=FALSE}
module_list <- list(
  "My First Data table" = dv.general::mod_data_table(
    dataset = dv.manager::mm_dispatch("filtered_dataset", "adsl"),
    module_id = "mod1"
  )
)
```

Please notice that in the dataset parameter we do not directly pass a dataset (e.g. a data.frame or a tibble), but just a string indicating which of the datasets we loaded before we want to display in this table. Please refer to these vignettes **\`vignette("how_does_work")\`** for the explanation on how data is dispatched to the different modules.

## Launching our first application

To launch our the application we now just need to call the run_app function in **dv.manager**.

```{r, eval=FALSE}
dv.manager::run_app(
  data = list("DS" = data),
  module_list = module_list,
  filter_data = "adsl"
)
```

