---
title: "logging"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{logging}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# How to use it

You can start using the logger simply by activating it and sending a message.

```{r, eval=FALSE}
dv.manager::log_activate()
rlang::inform("Logged message")
dv.manager::log_deactivate()
```

In its default state, the logger will capture and decorate any 'message' or 'warning' condition. Simply put,
any message you send using `message`, `warning`, `rlang::inform` and `rlang::warn`. Usually we will favour
the `rlang` functions as adding fields to use in our logger is simpler.

# Using logging levels

The logger can have different print granularities depending on the logging level selected. A different level can be specified in each
of the logging messages sent by including a logging level field in it.

```{r, eval=FALSE}
dv.manager::log_activate(dv.manager:::log_default_handlers(level = "info"))
rlang::inform("This will be printed", level = "info")
rlang::inform("This will not", level = "debug")
dv.manager::log_deactivate()
```

The logging levels and their numerical value can be checked with:

```{r, eval=FALSE}
dv.manager::log_get_levels()
```

# Logging messages sent from your module

A common use case will be using the logger to send messages from inside your module. To do this you can include the package name
as an argument when you send the message. This can be done in a simple way using `rlang`

```{r, eval=FALSE}
dv.manager::log_activate()
rlang::inform("An inform with package name", package = "my_package")
dv.manager::log_deactivate()
```

You can use the `packageName` function to dynamically get the name of your package.

You can also call `log_use_log` from `dv.manager` in the root of the package containing the module. This will add a file to your `R/`
directory with two convenience functions that automatically add the package name.

```{r, eval=FALSE}
# Will create an R/utils_logging.R file with two convenience functions
dv.manager::log_use_log()
devtools::load_all()
dv.manager::log_activate()
log_inform("An inform with package name")
log_warning("A warning with package name")
dv.manager::log_deactivate()
```

# Advanced logging

`dv.manager` logging system is based on [R's condition system](https://adv-r.hadley.nz/conditions.html).
This logging system simply specifies a set of `globalCallingHandlers` that will capture and log the messages,
warnings and/or errors as specified by the handlers.

## The default handlers

This is a low level logging and that makes it very flexible.

By default, the condition will contain the same fields as any standard `rlang::inform` or `rlang::warn` being the most important:

  - message: [character(1)] The message to be logged

It expects that the module developer has added the fields:

  - package: [character(1)] The name of the package that sent the condition (added automatically if `log_info` and `log_warn` were used).
  - level: [numeric(1)] a level for logging granularity.

The handlers will try to add the following fields to the condition before formatting:

  - date: [character(1)] A timestamp for the message.
  - ns: [character(1)] The namespace of the module that sent the condition.
  - sess_id: [character(1)] The Shiny session id that sent the condition.
  - short_sess_id: [character(1)] A truncated version of the sess_id to improve human-readibility.

The default handlers present and format this information.

```{r, eval = FALSE}
library(magrittr)

log_default_handlers <- function(level = 999) {
  format_str <- "[{date}][{package}|{short_sess_id}|{ns}]:{message}"

  cnd_to_str <- function(cnd) {
    cnd %>%
      dv.manager::log_add_date() %>%
      dv.manager::log_add_ns() %>%
      dv.manager::log_add_sess_id() %>%
      dv.manager::log_add_short_sess_id() %>%
      dv.manager::log_format(format_str)
  }

  list(
    message = function(cnd) {
      if (dv.manager::log_check_print(cnd, level)) cli::cli_alert_info(cnd_to_str(cnd))

      rlang::cnd_muffle(cnd)
    },
    warning = function(cnd) {
      if (dv.manager::log_check_print(cnd, level)) cli::cli_alert_warning(cnd_to_str(cnd))
      rlang::cnd_muffle(cnd)
    }
  )
}
```

The `log_add*` functions add extra fields to the condition that can be used later by `log_format()` to create a logging message. To print these messages we use the [`cli`](https://cli.r-lib.org/) package.
The `log_format` function is a convenience function that allows quick formatting of the condition. It allows using a [`glue`](https://glue.tidyverse.org/)-like string where the parameters are fields in the condition.
If a field used in the format is not present an `NA` value is returned instead.

## Creating custom handlers

The default handlers are convenient, but the logging behavior can be modified by using a different set of handlers...

```{r, eval=FALSE}
library(magrittr)

my_handlers <- function() {
  format_str <- "Date:{date}|Pkg:{package}|{custom_field}:{message}"

  cnd_to_str <- function(cnd) {
    cnd %>%
      dv.manager::log_add_date() %>%
      dv.manager::log_format(format_str)
  }

  list(
    message = function(cnd) {
      if (cnd[["print_me"]]) cli::cli_alert_info(cnd_to_str(cnd))
      rlang::cnd_muffle(cnd)
    },
    warning = function(cnd) {
      cli::cli_alert_warning(cnd_to_str(cnd))
      rlang::cnd_muffle(cnd)
    }
  )
}

dv.manager::log_activate(my_handlers())
dv.manager:::log_inform("I will print this and its custom_field will be 1", print_me = TRUE, custom_field = "1")
dv.manager:::log_inform("I will print this and its custom_field will be NA", print_me = TRUE)
dv.manager:::log_inform("I will not print this", print_me = FALSE)
dv.manager::log_deactivate()
```

As you can see we can add new fields for logging, we can supress some messages depending on the fields, etc. This allows filtering logging messages
by package, session, logging level, etc. So it adapts to our particular need during development.

# Troubleshooting

- When `log_activate` is called, an error similar to the following appears:

> Error in globalCallingHandlers(...) : should not be called with handlers on the stack

 The error above means you are activating the logger from an environment that already has defined some calling handlers,
 usually with `withCallingHandlers`. `shiny` in particular does this when an app starts, therefore you cannot activate logging once you are inside
 the application. The logger is based on setting `globalCallingHandlers` and setting those is tricky within other functions. The recommendation is to make the call to `log_activate` in your root
 session and not within functions. 
