---
title: "Passing data between modules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Passing data between modules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = "styler"
)
```

You can use the output of one module as the input of another module. 

In the following app you can select a value in one the modules and use its value in another module.

```{r, eval=FALSE}
com_test_UI <- function(id, choices = c(1, 2, 3), message) { # nolint
  ns <- shiny::NS(id)
  shiny::tagList(
    shiny::selectizeInput(ns("select"), label = "Select a number", choices = choices),
    shiny::p(message),
    shiny::textOutput(ns("output"))
  )
}

com_test_server <- function(id, value) {
  module <- function(input, output, session) {
    output$output <- shiny::renderText({
      shiny::req(value())
    })

    return(shiny::reactive(input$select))
  }

  return(
    shiny::moduleServer(
      id,
      module
    )
  )
}

mod_com_test <- function(choices, message, value, mod_id) {
  mod <- list(
    ui = function(id) {
      com_test_UI(id, choices, message)
    },
    server = function(server_args) {
      com_test_server(id = mod_id, value = server_args[["module_output"]]()[[value]])
    },
    module_id = mod_id
  )
  mod
}

dv.manager::run_app(
  data = list(),
  module_list = list(
    "Send and Receive 1" = mod_com_test(
      choices = 1:3,
      message = "The other module has selected",
      value = "mod_2",
      mod_id = "mod_1"
    ),
    "Send and Receive 2" = mod_com_test(
      choices = c("a", "b", "c"),
      message = "The other module has selected",
      value = "mod_1",
      mod_id = "mod_2"
    ),
    "Simple" = dv.manager::mod_simple("adsl", "modSimp")
  ),
  filter_data = "adsl",
  filter_key = "USUBJID"
)
```

Notice how in the mod_table wrapper we call `afmm[["module_output"]]()` by adding the two parenthesis at the end. In this case it is not because `afmm[["module_output"]]()` is a reactive object, we are calling it outside a reactive environment, but because it is a function that returns the list of values returned by the modules.

# IMPORTANT NOTE
If the element returned, by `afmm[["module_output"]]()[[selection]]` is not a reactive, it must be transformed into a reactive (usually a metaReactive) in the wrapper `metaReactive(afmm[["module_output"]]()[[selection]])`. This is done this way because at the moment the modules are invoked, module_output does not exist yet and its value should be retrieved after all the server functions have run and the variable exists inside the app_server environment.
