---
title: "Add timestamp to app data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Add timestamp to app data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Data timestamp

`dv.manager` shows the last time of modification of the data used in the application. To do this it checks the `meta` attribute of each of the datasets loaded in the app. Specifically,
it expects that `meta` is a list and that it has an entry `mtime` which contains a POSIXct date time.

```{r eval=FALSE}
data <- list(
  dummy = list(
    adsl = tibble::tibble(
      USUBJID = c(1, 2, 3),
      AGE = c(1, 2, 3)
    )
  )
)

# This attribute is automatically set when using dv.loader
attr(data[["dummy"]][["adsl"]], "meta") <- list(mtime = as.POSIXct("2023-02-28 12:00:00 CET"))

# This app shows the date specified in the previous step.
dv.manager::run_app(
  data = data,
  module_list = list(),
  filter_data = "adsl"
)
```

Usually, `dv.loader` sets this attribute automatically when loading the data and  no further action is needed. Nonetheless, when the data is modified after loading, this
attribute is deleted in some cases, see an example below, and it is necessary to keep track of it and reset it when needed.

```{r eval=FALSE}
data <- list(
  dummy = list(
    adsl = tibble::tibble(
      USUBJID = c(1, 2, 3),
      AGE = c(1, 2, 3)
    )
  )
)

# Normally set by dv.loader
attr(data[["dummy"]][["adsl"]], "meta") <- list(mtime = as.POSIXct("2023-02-28 12:00:00 CET"))

adsl2 <- tibble::tibble(
  USUBJID = c(1, 2, 3),
  COUNTRY = c("A", "B", "C")
)

# Save the date of the original dataset
adsl_date <- attr(data[["dummy"]][["adsl"]], "meta")

# Attribute is lost by the left_join
data[["dummy"]][["adsl"]] <- dplyr::left_join(adsl2, data[["dummy"]][["adsl"]])

# Reset date attribute afterwards
attr(data[["dummy"]][["adsl"]], "meta") <- adsl_date

dv.manager::run_app(
  data = data,
  module_list = list(),
  filter_data = "adsl"
)
```
